### 2019-06-08 files/20190608_ISSP09juitsa.Rmd
june06

---
title: ''
author: "Julio César Iturra Sanhueza - jciturra@uc.cl"
output: html_document
---

```{r}
Sys.setlocale("LC_ALL", 'en_US.UTF-8') # encoding
{
  library(haven)
  library(dplyr)
}
```